\chapter*{Acknowledgments}

%Insert your acknowledgments below the line
%-------------------------------------------

I want to begin by thanking David for introducing me to phase separation and the fascinating world of active droplets.
His constant words of support, patience and guidance made this four year journey enjoyable. 
He made sure he had time to answer my most trivial questions and made sure the group atmosphere remained conducive for research, which made the group a really warm place to be in.
Amongst many other things, he helped me realize the value of asking the right questions, how to make impactful presentations and how to see the big picture.
He was accommodating enough to have of endless hours of discussions with him, specially when it was just Jan and I in the group.
Thank you for everything David, you are an amazing supervisor!

I would also like to thank my TAC members, Prof. Stefan Klumpp and Prof. Michael Wilczek for their encouraging support, regular feedback, checking in on my credit progress and co-supervision.
Additionally, I again want to thank Prof. Stefan Klumpp for agreeing to review this thesis.
I want to thank Dr. Lucas Menou and (soon to be a Dr.) Jan for proofreading parts of this thesis. 

I want to thank Jan Kirschbaum for his four years of being a wonderful office mate, football companion, German translator, Indian food connoisseur, beer drinker, meme lover and much much more.
Thanks for making the office a lively place to be in, it will be missed. 
I want to specially thank him for his help in all the paperwork I had to do - from filling out travel expense reports, applying for conferences etc.
I want to thank the entire Zwicker group (in no particular order) - Noah, Tefa, Marcel, Swati, Lucas, Johannes, Malte, Pam and Manish for making these years wonderful and warm.
I will miss the amazing atmosphere you guys had in the group.
Hansol, Antoine, Olinka and Swati Kaushik - I will miss our impromptu trips, Vapiano dinners, hikes and weekend beer sessions.
A special thanks to other people at the institute (in no particular order) - Shubhadeep, Puneet, Prashanth, Babak, Alex, Rodrigo, Kris, Monika, Akirnori, Antonio who made my time memorable and enjoyable.

Now I would like to thank Shaunak Phatak, Jaydeep Deshpande, Kaustubh Deshpande, Ajinkya Bhave, Abhishek Deshpande, Heramb Karnataki, Anay Joshi and Omkar Ghalsasi for over ten years of being together and managed to stay close, despite being literally all over the world.

I am sincerely grateful to Prof. Mahesh and Prof. Sumesh at IIT Madras, who introduced to me the world of research, the value of asking the right questions and pushing oneself further.
``If I have seen further it is by standing on the shoulders of giants", and they were the giants who led me into this world, believed in me and stood by me from $2014-2017$.
Thanks you sirs, I am deeply indebted.

I am deeply grateful to my wife Radhika, for being with me through \textit{everything}.
I thank her for her company, sometimes just being there, witnessing my highs and lows, my occasional rants and listening about my Ph.D work for hours.
A special thanks to her for keeping me sane during these last weeks of thesis submission. 
A special mention should go to your carefully picked inshorts news snippets, which made my days brighter regardless of the weather. 
A special thanks to Shalmali, with whom I shared many interesting discussions, and more importantly, memes and stickers. 

To Aai and Chaitanya - thank for your unconditional love, patience and support throughout my journey.
I am grateful to Aai and Baba for bringing us up in unconventionally difficult circumstances, and pushing me to achieve what I want. 
Thank you Baba, who inspired me to take up this journey and for being able to do what few people can. 
You helped me pursue what I felt was impossible.
I dedicate this work to you.